---
title: "PRISMA_flow_diagram_ns"
output: html_document
date: "2026-09-25"
---

```{r setup, include=FALSE}
# Load packages
library(readxl)
library(readr)
library(tidyverse)
```

# Step 1: Population
```{r}
# List of all articles, including whether the term "pre*reg*" was found in the full text and whether an OSF link was found
fulltext_unfiltered <- read_csv("fulltext_unfiltered.csv") 
  
fulltext_unfiltered |> nrow()
# Total number of articles in the population: 3317
```

# Step 2: Initial (automated) screening
Searched texts for instances of *"preregistered", "pre-registered",
  "preregistration", "pre-registration"* and extracted OSF link for articles that had one.
```{r}
fulltext_unfiltered |> 
  filter(prereg_found & osf_link_found) |> 
  nrow()
# Potential ellgible articles: 1138

# Did not include any “pre*regist*” variant 
fulltext_unfiltered |> 
  filter(prereg_found == F) |> 
  nrow()
# n = 2087

#Included “pre*regist*” but did not contain OSF links in paper 
fulltext_unfiltered |> 
  filter(prereg_found == T & osf_link_found == F) |> 
  nrow()
# n = 92
```

# Step 3: LLM classifier screening
We built an LLM classifier with the gpt-5-nano model (see preregistration_classifier.ipynb) to identify which articles that contained one of keywords and an OSF link actually contained preregistered studies, and if so, which studies were preregistered and the OSF link provided. 
```{r}
prereg_classified <- read_csv("prereg_classified.csv", col_select = -fulltext)

prereg_classified_JEPLMC <- read_csv("prereg_classified_JEPLMC.csv", col_select = -fulltext) # this journal was originally labelled incorrectly, so had to be processed by the LLM in a second batch

# merge them together
prereg_classified <- bind_rows(prereg_classified, prereg_classified_JEPLMC)

prereg_classified |> distinct(uid) |> nrow()
# Total (check this is consisntent with n elligble articles above): 1036 [consistent]

prereg_classified |> 
  group_by(uid) |> 
  summarise(
    any_na_real_prereg = any(is.na(real_prereg)),
    any_true_real_prereg = any(real_prereg == TRUE, na.rm = TRUE),
    any_linked = any(real_prereg == TRUE & !is.na(uses_osf_template), na.rm = TRUE)
  ) |> 
  mutate(category = case_when(
    any_na_real_prereg   ~ "llm_could_not_parse",
    !any_true_real_prereg ~ "not_preregistered",
    any_linked            ~ "at_least_one_prereg",
    TRUE                  ~ "llm_could_not_link_to_prereg"
  )) |> 
  count(category)

# at_least_one_prereg = 948

# --
# not_preregistered =	98	
# llm_could_not_link_to_prereg = 91			
# llm_could_not_parse = 1			
```
# Step 4: Uses OSF's Preregistration Template
We used OSF’s API to determine the template of that preregistration and the direct link to the preregistration (authors often link to the repository, rather than to the specific registration).
```{r}
prereg_classified |> 
  filter(real_prereg & !is.na(uses_osf_template)) |> 
  group_by(uid) |> 
  summarise(any_uses_osf_template = any(uses_osf_template)) |> 
  mutate(category = case_when(
    any_uses_osf_template ~ "any_uses_osf_template",
    TRUE ~ "none_use_osf_template"
  )) |> 
  count(category)

# any_uses_osf_template	= 230			
# none_use_osf_template	= 718	
```

# Step 5: Final Sample
```{r}
MASTER_list <- read_csv("feed_to_regcheck/MASTER_list.csv")

# final coded
MASTER_list |> 
  filter(is.na(exclude)) |> 
  nrow()

# n articles = 100

# n excluded
MASTER_list |> 
  filter(exclude == "Excluded") |> 
  nrow()

# n articles = 14

# n not coded
template_uids <- prereg_classified |> 
  filter(real_prereg & uses_osf_template) |> 
  group_by(DOI) |> 
  summarise(any_uses_osf_template = any(uses_osf_template)) |> 
  filter(any_uses_osf_template) |> 
  distinct(DOI)

# n not randomly selected
template_uids |> 
  anti_join(MASTER_list, by = "DOI") |> 
  nrow()
# 116 articles not randomly selected
```

