---
title: "Untitled"
format: html
editor: visual
---

# Import libraries

```{r}

library(MASS)
library(tidyverse)
library(lme4)

```

## Define sorted pair helper for raters

```{r}

make_sorted_rater_pair <- function(rater1_id, rater2_id) {
  map2_chr(
    rater1_id,
    rater2_id,
    ~ paste(sort(c(.x, .y)), collapse = "__")
  )
}

```

## Define data-generating function

```{r}

simulate_regcheck_study_data <- function(
    number_of_papers = 100,
    number_of_raters = 5,
    number_of_raters_per_paper = 2,
    number_of_dimensions = 9,
    intercept_agreement_logit = 1.4,
    agreement_comparison_effect_logit = -0.3,
    intercept_cosine_logit = 1.8,
    cosine_comparison_effect_logit = -0.2,
    latent_outcome_correlation = 0.5,
    random_effect_rater = 0.4,
    random_effect_dimension = 0.3,
    random_effect_paper = 0.5,
    seed = NULL
) {

  # basic checks 
  if (!is.null(seed)) {
    set.seed(seed)
  }

  if (number_of_raters_per_paper < 2) {
    stop("number_of_raters_per_paper must be at least 2.")
  }

  if (number_of_raters_per_paper > number_of_raters) {
    stop("number_of_raters_per_paper cannot exceed number_of_raters.")
  }

  # assign rater ids and regcheck ids
  human_rater_ids <- paste0("human_", seq_len(number_of_raters))
  regcheck_id <- "RegCheck"

  # create paper dimension data
  paper_ids <- paste0("paper_", seq_len(number_of_papers))
  dimension_ids <- paste0("dimension_", seq_len(number_of_dimensions))
  all_rater_ids <- c(human_rater_ids, regcheck_id)

  # specify paper random effect
  paper_random_effects <- tibble(
    paper_id = paper_ids,
    paper_random_effect_agreement = rnorm(
      n = number_of_papers,
      mean = 0,
      sd = random_effect_paper
    ),
    paper_random_effect_cosine = rnorm(
      n = number_of_papers,
      mean = 0,
      sd = random_effect_paper
    )
  )

  # specify dimension random effect
  dimension_random_effects <- tibble(
    dimension = dimension_ids,
    dimension_random_effect_agreement = rnorm(
      n = number_of_dimensions,
      mean = 0,
      sd = random_effect_dimension
    ),
    dimension_random_effect_cosine = rnorm(
      n = number_of_dimensions,
      mean = 0,
      sd = random_effect_dimension
    )
  )

  # define all pairs
  all_possible_rater_pairs <- combn(
    x = all_rater_ids,
    m = 2,
    simplify = FALSE
  ) |>
    map_chr(~ paste(sort(.x), collapse = "__"))

  # define random effects for raters
  rater_pair_random_effects <- tibble(
    rater_pair = all_possible_rater_pairs,
    rater_pair_random_effect_agreement = rnorm(
      n = length(all_possible_rater_pairs),
      mean = 0,
      sd = random_effect_rater
    ),
    rater_pair_random_effect_cosine = rnorm(
      n = length(all_possible_rater_pairs),
      mean = 0,
      sd = random_effect_rater
    )
  )

  # randomly select raters for each paper
  selected_raters_by_paper <- tibble(
    paper_id = paper_ids,
    selected_human_raters = map(
      seq_len(number_of_papers),
      ~ sample(
        x = human_rater_ids,
        size = number_of_raters_per_paper,
        replace = FALSE
      )
    )
  )

  # create final study design (incl. specified rater assignment)
  study_design <- crossing(
    paper_id = paper_ids,
    dimension = dimension_ids
  ) |>
    left_join(selected_raters_by_paper, by = "paper_id") |>
    mutate(
      comparison_rows = map(
        selected_human_raters,
        function(selected_human_raters) {

          human_human_pairs <- combn(
            x = selected_human_raters,
            m = 2,
            simplify = FALSE
          )

          human_human_rows <- map_dfr(
            human_human_pairs,
            function(current_pair) {
              tibble(
                comparison_label = "human_human",
                comparison_type = 0,
                rater1_id = current_pair[1],
                rater2_id = current_pair[2]
              )
            }
          )

          ai_human_rows <- tibble(
            comparison_label = "ai_human",
            comparison_type = 1,
            rater1_id = regcheck_id,
            rater2_id = selected_human_raters
          )

          bind_rows(human_human_rows, ai_human_rows)
        }
      )
    ) |>
    select(-selected_human_raters) |>
    unnest(comparison_rows) |>
    mutate(
      rater_pair = make_sorted_rater_pair(rater1_id, rater2_id)
    )

  # add random effects to data
  simulated_data <- study_design |>
    left_join(paper_random_effects, by = "paper_id") |>
    left_join(dimension_random_effects, by = "dimension") |>
    left_join(rater_pair_random_effects, by = "rater_pair")

  # specify covariance between cosine and agreement
  latent_error_correlation_matrix <- matrix(
    c(
      1, latent_outcome_correlation,
      latent_outcome_correlation, 1
    ),
    nrow = 2,
    byrow = TRUE
  )

  # simulate latent correlation 
  latent_errors <- MASS::mvrnorm(
    n = nrow(simulated_data),
    mu = c(0, 0),
    Sigma = latent_error_correlation_matrix
  )

  # add latent correlations to data, simulate data
  simulated_data <- simulated_data |>
    mutate(
      latent_error_agreement = latent_errors[, 1],
      latent_error_cosine = latent_errors[, 2]
    ) |>
    # specify linear predictions for each outcome and outcomes themselves
    mutate(
      linear_predictor_agreement =
        intercept_agreement_logit +
        agreement_comparison_effect_logit * comparison_type +
        paper_random_effect_agreement +
        dimension_random_effect_agreement +
        rater_pair_random_effect_agreement,
      
      # below lines: render agreement judgement
      probability_of_agreement = plogis(linear_predictor_agreement),
      
      # threshold needs to invert 
      agreement_threshold = qnorm(1 - probability_of_agreement),
      agree = as.integer(
        latent_error_agreement > agreement_threshold
        )
    ) |>
    mutate(
      linear_predictor_cosine =
        intercept_cosine_logit +
        cosine_comparison_effect_logit * comparison_type +
        paper_random_effect_cosine +
        dimension_random_effect_cosine +
        rater_pair_random_effect_cosine +
        latent_error_cosine,
       # render cosine similarity 
      cosine_similarity = plogis(linear_predictor_cosine)
    )

  simulated_data
}


```

## Define analysis function

```{r}

analyse_regcheck_study_data <- function(
    simulated_data,
    confidence_level = 0.95,
    number_of_parameter_draws = 2000
) {

  # RQ1 model
  agreement_model <- glmer(
    agree ~ comparison_type +
      (1 | paper_id) +
      (1 | dimension) +
      (1 | rater_pair),
    family = binomial,
    data = simulated_data
  )

  # RQ2 model
  cosine_model <- lmer(
    cosine_similarity ~ comparison_type +
      (1 | paper_id) +
      (1 | dimension) +
      (1 | rater_pair),
    data = simulated_data
  )

  # define alpha and quantiles
  alpha <- 1 - confidence_level
  lower_quantile <- alpha / 2
  upper_quantile <- 1 - alpha / 2

  human_human_data <- simulated_data |>
    filter(comparison_type == 0)

  # specify noninferiority margins
  agreement_noninferiority_margin <- 0.5 * sd(human_human_data$agree)
  cosine_noninferiority_margin <- 0.5 * sd(human_human_data$cosine_similarity)

  # get fixed effect estimates
  agreement_fixed_effect_estimates <- fixef(agreement_model)
  agreement_fixed_effect_vcov <- as.matrix(vcov(agreement_model))

  agreement_parameter_draws <- MASS::mvrnorm(
    n = number_of_parameter_draws,
    mu = agreement_fixed_effect_estimates,
    Sigma = agreement_fixed_effect_vcov
  )

  agreement_intercept_draws <- agreement_parameter_draws[, "(Intercept)"]
  agreement_comparison_draws <- agreement_parameter_draws[, "comparison_type"]

  predicted_human_human_agreement_draws <- plogis(agreement_intercept_draws)
  predicted_ai_human_agreement_draws <- plogis(
    agreement_intercept_draws + agreement_comparison_draws
  )

  agreement_difference_draws <-
    predicted_ai_human_agreement_draws -
    predicted_human_human_agreement_draws

  agreement_difference_estimate <- mean(agreement_difference_draws)

  agreement_difference_confidence_interval <- quantile(
    agreement_difference_draws,
    probs = c(lower_quantile, upper_quantile)
  )

  agreement_noninferior <-
    agreement_difference_confidence_interval[1] >
    (-agreement_noninferiority_margin)

  cosine_fixed_effects <- summary(cosine_model)$coefficients

  cosine_difference_estimate <- cosine_fixed_effects["comparison_type", "Estimate"]
  cosine_difference_standard_error <- cosine_fixed_effects["comparison_type", "Std. Error"]

  z_critical_value <- qnorm(1 - alpha / 2)

  cosine_difference_confidence_interval <- c(
    cosine_difference_estimate -
      z_critical_value * cosine_difference_standard_error,
    cosine_difference_estimate +
      z_critical_value * cosine_difference_standard_error
  )

  cosine_noninferior <-
    cosine_difference_confidence_interval[1] >
    (-cosine_noninferiority_margin)

  tibble(
    agreement_noninferiority_margin = agreement_noninferiority_margin,
    agreement_difference_estimate = agreement_difference_estimate,
    agreement_confidence_interval_lower = agreement_difference_confidence_interval[1],
    agreement_confidence_interval_upper = agreement_difference_confidence_interval[2],
    agreement_noninferior = agreement_noninferior,
    cosine_noninferiority_margin = cosine_noninferiority_margin,
    cosine_difference_estimate = cosine_difference_estimate,
    cosine_confidence_interval_lower = cosine_difference_confidence_interval[1],
    cosine_confidence_interval_upper = cosine_difference_confidence_interval[2],
    cosine_noninferior = cosine_noninferior
  )
}

```

## Define function runners

for single run (test) and full run

```{r}

run_one_regcheck_simulation <- function(
    number_of_papers = 100,
    number_of_raters = 5,
    number_of_raters_per_paper = 2,
    number_of_dimensions = 9,
    intercept_agreement_logit = 0,
    agreement_comparison_effect_logit = -0.3,
    intercept_cosine_logit = 1,
    cosine_comparison_effect_logit = -0.2,
    latent_outcome_correlation = 0.5,
    random_effect_rater = 0.4,
    random_effect_dimension = 0.3,
    random_effect_paper = 0.5,
    seed = NULL
) {

  simulated_data <- simulate_regcheck_study_data(
    number_of_papers = number_of_papers,
    number_of_raters = number_of_raters,
    number_of_raters_per_paper = number_of_raters_per_paper,
    number_of_dimensions = number_of_dimensions,
    intercept_agreement_logit = intercept_agreement_logit,
    agreement_comparison_effect_logit = agreement_comparison_effect_logit,
    intercept_cosine_logit = intercept_cosine_logit,
    cosine_comparison_effect_logit = cosine_comparison_effect_logit,
    latent_outcome_correlation = latent_outcome_correlation,
    random_effect_rater = random_effect_rater,
    random_effect_dimension = random_effect_dimension,
    random_effect_paper = random_effect_paper,
    seed = seed
  )

  analysis_results <- analyse_regcheck_study_data(simulated_data)

  list(
    simulated_data = simulated_data,
    analysis_results = analysis_results
  )
}


run_many_regcheck_simulations <- function(
    number_of_simulations = 100,
    number_of_papers = 100,
    number_of_raters = 5,
    number_of_raters_per_paper = 2,
    number_of_dimensions = 9,
    intercept_agreement_logit = 1.4,
    agreement_comparison_effect_logit = -0.3,
    intercept_cosine_logit = 1.8,
    cosine_comparison_effect_logit = -0.2,
    latent_outcome_correlation = 0.5,
    random_effect_rater = 0.4,
    random_effect_dimension = 0.3,
    random_effect_paper = 0.5,
    seed = 123
) {

  map_dfr(
    seq_len(number_of_simulations),
    function(simulation_index) {

      current_seed <- seed + simulation_index - 1

      one_simulation <- tryCatch(
        run_one_regcheck_simulation(
          number_of_papers = number_of_papers,
          number_of_raters = number_of_raters,
          number_of_raters_per_paper = number_of_raters_per_paper,
          number_of_dimensions = number_of_dimensions,
          intercept_agreement_logit = intercept_agreement_logit,
          agreement_comparison_effect_logit = agreement_comparison_effect_logit,
          intercept_cosine_logit = intercept_cosine_logit,
          cosine_comparison_effect_logit = cosine_comparison_effect_logit,
          latent_outcome_correlation = latent_outcome_correlation,
          random_effect_rater = random_effect_rater,
          random_effect_dimension = random_effect_dimension,
          random_effect_paper = random_effect_paper,
          seed = current_seed
        ),
        error = function(error_message) {
          NULL
        }
      )

      if (is.null(one_simulation)) {
        return(
          tibble(
            simulation_index = simulation_index,
            model_converged = FALSE,
            agreement_noninferiority_margin = NA_real_,
            agreement_difference_estimate = NA_real_,
            agreement_confidence_interval_lower = NA_real_,
            agreement_confidence_interval_upper = NA_real_,
            agreement_noninferior = NA,
            cosine_noninferiority_margin = NA_real_,
            cosine_difference_estimate = NA_real_,
            cosine_confidence_interval_lower = NA_real_,
            cosine_confidence_interval_upper = NA_real_,
            cosine_noninferior = NA
          )
        )
      }

      one_simulation$analysis_results |>
        mutate(
          simulation_index = simulation_index,
          model_converged = TRUE,
          .before = 1
        )
    }
  )
}

```

### Define margin check function

```{r}

classify_margin_position <- function(true_deficit, average_margin, boundary_buffer = 0.10) {

  lower_boundary <- average_margin * (1 - boundary_buffer)
  upper_boundary <- average_margin * (1 + boundary_buffer)

  case_when(
    true_deficit < lower_boundary ~ "inside_margin",
    true_deficit > upper_boundary ~ "outside_margin",
    TRUE ~ "near_margin"
  )
}


```

## Run single simulation (test)

```{r}

one_simulation <- run_one_regcheck_simulation(
  number_of_papers = 100,
  number_of_raters = 5,
  number_of_raters_per_paper = 2,
  intercept_agreement_logit = qlogis(0.8),
  agreement_comparison_effect_logit = qlogis(0.75) - qlogis(0.8),
  intercept_cosine_logit = qlogis(0.6),
  cosine_comparison_effect_logit = qlogis(0.5) - qlogis(0.6),
  latent_outcome_correlation = 0.7,
  random_effect_rater = 0.30,
  random_effect_dimension = 0.25,
  random_effect_paper = 0.40,
  seed = 123
)

one_simulation$simulated_data |>
  count(comparison_label)

# check correlation is roughly rocovered since I had some difficulty with this
cor(one_simulation$simulated_data$agree, 
    one_simulation$simulated_data$cosine_similarity)

one_simulation$simulated_data |>
  count(rater_pair, sort = TRUE)


```

## Define simulation experiment parameters

```{r}

scenario_parameters <- tribble(
  ~scenario_label,
  ~human_human_agreement,
  ~ai_human_agreement,
  ~human_human_cosine,
  ~ai_human_cosine,
  "high_and_noninferior", 0.80, 0.75, 0.60, 0.55,
  "high_and_inferior",    0.80, 0.70, 0.60, 0.50,
  "low_and_noninferior",  0.70, 0.65, 0.40, 0.35,
  "low_and_noninferior", 0.70, 0.60, 0.40, 0.30
) |>
  mutate(
    intercept_agreement_logit = qlogis(human_human_agreement),
    agreement_comparison_effect_logit =
      qlogis(ai_human_agreement) - qlogis(human_human_agreement),
    intercept_cosine_logit = qlogis(human_human_cosine),
    cosine_comparison_effect_logit =
      qlogis(ai_human_cosine) - qlogis(human_human_cosine),
    true_agreement_deficit = human_human_agreement - ai_human_agreement,
    true_cosine_deficit = human_human_cosine - ai_human_cosine
  )

design_parameters <- crossing(
  number_of_papers = c(50, 100, 150),
  number_of_raters = c(5, 7),
  number_of_raters_per_paper = c(2, 3),
  latent_outcome_correlation = c(0.6, 0.8)
)

random_effect_parameters <- tibble(
  random_effect_rater = 0.30,
  random_effect_dimension = 0.25,
  random_effect_paper = 0.40
)

# final parameter grid
parameter_grid <- crossing(
  design_parameters,
  scenario_parameters,
  random_effect_parameters
) |>
  mutate(
    number_of_simulations = 1000,
    seed = 123
  )

```

## Run full simulation

```{r}

simulation_results <- parameter_grid |>
  mutate(
    simulation_output = pmap(
      .l = list(
        number_of_simulations = number_of_simulations,
        number_of_papers = number_of_papers,
        number_of_raters = number_of_raters,
        number_of_raters_per_paper = number_of_raters_per_paper,
        intercept_agreement_logit = intercept_agreement_logit,
        agreement_comparison_effect_logit = agreement_comparison_effect_logit,
        intercept_cosine_logit = intercept_cosine_logit,
        cosine_comparison_effect_logit = cosine_comparison_effect_logit,
        latent_outcome_correlation = latent_outcome_correlation,
        random_effect_rater = random_effect_rater,
        random_effect_dimension = random_effect_dimension,
        random_effect_paper = random_effect_paper,
        seed = seed
      ),
      .f = run_many_regcheck_simulations
    )
  )

write_csv("results/simulation_results.csv")

simulation_results_long <- simulation_results |>
  unnest(simulation_output)

```

## Summarise simulation results

```{r}

simulation_summary <- simulation_results_long |>
  group_by(
    scenario_label,
    number_of_papers,
    number_of_raters,
    number_of_raters_per_paper,
    latent_outcome_correlation,
    human_human_agreement,
    ai_human_agreement,
    human_human_cosine,
    ai_human_cosine,
    true_agreement_deficit,
    true_cosine_deficit
  ) |>
  summarise(
    proportion_converged = mean(model_converged, na.rm = TRUE),

    mean_agreement_margin = mean(agreement_noninferiority_margin, na.rm = TRUE),
    mean_cosine_margin = mean(cosine_noninferiority_margin, na.rm = TRUE),

    agreement_noninferiority_rate = mean(agreement_noninferior, na.rm = TRUE),
    cosine_noninferiority_rate = mean(cosine_noninferior, na.rm = TRUE),
    joint_noninferiority_rate = mean(
      agreement_noninferior & cosine_noninferior,
      na.rm = TRUE
    ),

    .groups = "drop"
  ) |>
  mutate(
    agreement_margin_position = classify_margin_position(
      true_deficit = true_agreement_deficit,
      average_margin = mean_agreement_margin
    ),
    cosine_margin_position = classify_margin_position(
      true_deficit = true_cosine_deficit,
      average_margin = mean_cosine_margin
    ),
    joint_margin_position = case_when(
      agreement_margin_position == "inside_margin" &
        cosine_margin_position == "inside_margin" ~ "inside_margin",
      agreement_margin_position == "outside_margin" |
        cosine_margin_position == "outside_margin" ~ "outside_margin",
      TRUE ~ "near_margin"
    )
  )

```

## Look specifically overall at number of papers

```{r}

number_of_papers_overall_effect <- simulation_summary |>
  group_by(
    number_of_papers,
    scenario_label,
    joint_margin_position
  ) |>
  summarise(
    mean_joint_noninferiority_rate = mean(joint_noninferiority_rate, na.rm = TRUE),
    mean_agreement_noninferiority_rate = mean(agreement_noninferiority_rate, na.rm = TRUE),
    mean_cosine_noninferiority_rate = mean(cosine_noninferiority_rate, na.rm = TRUE),
    mean_convergence_rate = mean(proportion_converged, na.rm = TRUE),
    .groups = "drop"
  )

number_of_papers_overall_effect

paper_count_decision_table <- number_of_papers_overall_effect |>
  pivot_wider(
    names_from = joint_margin_position,
    values_from = c(
      mean_joint_noninferiority_rate,
      mean_agreement_noninferiority_rate,
      mean_cosine_noninferiority_rate
    )
  )

paper_count_decision_table



```

## Plot effect of papers

```{r}

plot_number_of_papers <- number_of_papers_overall_effect |>
  mutate(
    joint_margin_position = factor(
      joint_margin_position,
      levels = c("inside_margin", "near_margin", "outside_margin")
    )
  ) |>
  pivot_longer(
    cols = c(
      mean_joint_noninferiority_rate,
      mean_agreement_noninferiority_rate,
      mean_cosine_noninferiority_rate
    ),
    names_to = "outcome_type",
    values_to = "noninferiority_rate"
  ) |>
  mutate(
    outcome_type = recode(
      outcome_type,
      mean_joint_noninferiority_rate = "Joint",
      mean_agreement_noninferiority_rate = "Agreement",
      mean_cosine_noninferiority_rate = "Cosine"
    )
  ) |>
  ggplot(
    aes(
      x = factor(number_of_papers),
      y = noninferiority_rate,
      group = joint_margin_position,
      linetype = joint_margin_position,
      shape = joint_margin_position
    )
  ) +
  geom_hline(
    yintercept = c(0.05, 0.80),
    linewidth = 0.4,
    alpha = 0.5
  ) +
  geom_point(size = 2.5) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ outcome_type * scenario_label, nrow = 3) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.1)
  ) +
  labs(
    x = "Number of papers",
    y = "Non-inferiority conclusion rate",
    linetype = "True position relative to margin",
    shape = "True position relative to margin",
    title = "Operating characteristics by number of papers",
    subtitle = "Rates of concluding non-inferiority when the true deficit is inside, near, or outside the margin"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

plot_number_of_papers


```
