---
title: "regcheck validation analysis plan"
output: html_document
date: "2026-03-10"
---

```{r setup, include=FALSE}
# Load packages
library(readr)
library(tidyverse)
library(knitr)
library(kableExtra)

library(krippendorffsalpha)
library(snow)
library(httr2) # for calculating cosine similarity

library(lme4)
library(emmeans)

# Read in data
data <- read_csv("test_pilot_data.csv", locale = locale(encoding = "latin1"))

options(scipen = 999) #turn off scientific notation
options(digits = 2)
```

# Initial cleaning of pilot data that we shouldn't need on the real data
```{r}
data <- data %>% 
  filter(doi != "10.3758/s13423-017-1395-4") %>% # filter out a paper that only one person coded
  mutate(
    measurvars_prereg = pmap_chr(
      select(., paste0("measurvars_prereg_", 4:13)),
      ~ { vals <- c(...); paste(vals[!is.na(vals) & vals != ""], collapse = " | ") }
    ),
    measurvars_paper = pmap_chr(
      select(., paste0("measurvars_paper_", 4:13)),
      ~ { vals <- c(...); paste(vals[!is.na(vals) & vals != ""], collapse = " | ") }
    ),
    measurvars_concerns = measurvars_serious  # rename for consistency with other seriousness variables
  )
```


# Step 1: Pivot the data wider
```{r}
data_long <- data %>%
  select(
    doi, coder, study_num,
    matches("^(hypothesis|datas|inex|samplesize|manvars|measurvars|stats|transf|missd)_(prereg|paper|notreg|deviation|concerns|conf|comments)$")
  ) %>%
  pivot_longer(
    cols = -c(doi, coder, study_num),
    names_to  = c("dimension", ".value"),
    names_pattern = "^(.+?)_(prereg|paper|notreg|deviation|concerns|conf|comments)$"
  )
```

# Step 2: Recode deviation decision into numeric value
```{r }
# Turn deviation decision into numeric value
data_long <- data_long %>%
  mutate(deviation = case_when(
    deviation == "No"  ~ 0L,
    deviation == "Yes" ~ 2L,
    str_detect(deviation, regex("vague|too vague|not reported", ignore_case = TRUE)) ~ 1L,
    TRUE ~ NA_integer_
  ))

# Define coder variables — update these if coder initials change
rc_coder     <- "RC"
human_coders <- sort(setdiff(unique(data_long$coder), rc_coder))
```

# Step 3: Convert to pairwise comparisons
```{r}
# Get all unique coder pairs
coders <- unique(data_long$coder)
coder_pairs <- combn(sort(coders), 2, simplify = FALSE)  # sort alphabetically (so that we don't get both BC_JC and JC_BC later)

# Build one pairwise row per doi x dimension x pair
data_pairs <- map_dfr(coder_pairs, function(pair) {
  c1 <- pair[1]; c2 <- pair[2]
  
  left  <- data_long %>% filter(coder == c1) %>%
    select(doi, study_num, dimension, 
           prereg_c1 = prereg, paper_c1 = paper,
           deviation_c1 = deviation, concerns_c1 = concerns)
  
  right <- data_long %>% filter(coder == c2) %>%
    select(doi, dimension,
           prereg_c2 = prereg, paper_c2 = paper,
           deviation_c2 = deviation, concerns_c2 = concerns)
  
  left %>%
    inner_join(right, by = c("doi", "dimension")) %>%
    mutate(
      coder1 = c1,
      coder2 = c2,
      coder_pair       = paste(c1, c2, sep = "_"),
      comparison_type  = if_else(str_detect(coder_pair, "RC"), "rc_human", "human_human"),
      deviation_agree  = as.integer(deviation_c1 == deviation_c2)
    )
})

# For computing cosine similarity, we have two relevant columns: i) text from the prereg and ii) text from the paper.
 # Let's compute cosine for each occasion, and treat this (the "source" of the quote) as a random effect.
data_pairs_cosine <- data_pairs %>%
  select(doi, study_num, dimension, coder1, coder2, coder_pair, 
         comparison_type, deviation_agree,
         prereg_c1, prereg_c2, paper_c1, paper_c2) %>%
  pivot_longer(
    cols = c(prereg_c1, prereg_c2, paper_c1, paper_c2),
    names_to = c("source", "coder_num"),
    names_pattern = "^(prereg|paper)_(c1|c2)$"
  ) %>%
  pivot_wider(names_from = coder_num, values_from = value) %>%
  rename(text_c1 = c1, text_c2 = c2)

data_pairs_agree <- data_pairs # We will use the non long form for the agree variable
```

# Step 4: Compute cosine similarity
```{r}
# --- HELPERS ---
# Helper: get embeddings for a character vector (single API call)
get_embeddings <- function(texts, 
                           model = "text-embedding-3-large",
                           api_key = Sys.getenv("OPENAI_API_KEY")) {
  resp <- request("https://api.openai.com/v1/embeddings") %>%
    req_auth_bearer_token(api_key) %>%
    req_body_json(list(input = as.list(texts), model = model)) %>%
    req_perform()
  
  resp_body_json(resp)$data %>%
    map(~ unlist(.x$embedding))
}

# Helper: cosine similarity between two numeric vectors
cosine_sim <- function(a, b) sum(a * b) / (sqrt(sum(a^2)) * sqrt(sum(b^2)))

# --- RUN FUNCTIONS ---
# Collect all unique non-empty texts across both columns and both coders
all_texts <- data_pairs %>%
  select(prereg_c1, prereg_c2, paper_c1, paper_c2) %>%
  unlist(use.names = FALSE) %>%
  { .[!is.na(.) & str_trim(.) != ""] } %>%
  unique()

# Single API call for all unique texts
embeddings <- get_embeddings(all_texts)
names(embeddings) <- all_texts

lookup <- function(text) embeddings[[text]]

data_pairs_cosine <- data_pairs_cosine %>%
  mutate(
    cosine = map2_dbl(text_c1, text_c2, function(a, b) {
      if (is.na(a) || a == "" || is.na(b) || b == "") return(NA_real_)
      cosine_sim(lookup(a), lookup(b))
    })
  ) %>% 
  select(-deviation_agree) # this will be duplicated, so let's remove it.

# Ensure columns are the right type
data_pairs_cosine <- data_pairs_cosine %>%
  mutate(comparison_type = factor(comparison_type, levels = c("human_human", "rc_human"))) 

data_pairs_agree <- data_pairs_agree %>%
  mutate(
    comparison_type = factor(comparison_type, levels = c("human_human", "rc_human")),
    deviation_agree = as.integer(deviation_agree)
  )
```

# Analysis
# RQ1: Cosine similarity
Descriptives
```{r}
data_pairs_cosine %>% 
  group_by(comparison_type) %>% 
  summarise(
    mean_cosine = mean(cosine, na.rm = TRUE),
    sd_cosine = sd(cosine, na.rm = TRUE)
  )
```

Inferential
```{r}
# Build the mlm
m_rq1 <- lmer(
  cosine ~ comparison_type + source + (1 | doi) + (1 | dimension) + (1 | coder_pair) + (1 |  doi:dimension:source),
  data = data_pairs_cosine
)

# View the model output
summary(m_rq1)

# Non-inferiority check
# criterion: 0.4 x SD of human-human cosine similarity
ni_criterion <- data_pairs_cosine %>%
  filter(comparison_type == "human_human") %>%
  pull(cosine) %>%
  sd(na.rm = TRUE) %>%
  { . * 0.4 }

# Contrast: rc_human vs human_human
contrast_rq1 <- emmeans(m_rq1, ~ comparison_type) %>%
  contrast("pairwise") %>%
  confint()

contrast_rq1

# Non-inferiority check: lower CI must be above -ni_criterion
# (i.e. RegCheck is not meaningfully worse than human-human)
contrast_rq1$lower.CL > -ni_criterion
```

# RQ1a -- cosine similarity by dimension (descriptive)
```{r}
rq1a <- data_pairs_cosine %>%
  filter(!is.na(cosine)) %>%
  summarise(
    mean_cosine = mean(cosine),
    sd_cosine  = sd(cosine),
    .by = c(dimension, comparison_type)
  ) %>%
  arrange(dimension, comparison_type)

rq1a
```

# RQ2: Deviation agreement
```{r}
# Build mlm
m_rq2 <- glmer(
  deviation_agree ~ comparison_type + (1 | doi) + (1 | dimension) + (1 | coder_pair) +  (1 | doi:dimension),
  data   = data_pairs_agree,
  family = binomial(link = "logit")
)

# View the model output
summary(m_rq2)

# Non-inferiority test
  # criterion: 0.2 x SD of human-human cosine similarity
hh_agree <- data_pairs_agree %>%
  filter(comparison_type == "human_human", !is.na(deviation_agree)) %>%
  pull(deviation_agree)

# Small epsilon to avoid qlogis(0) = -Inf and qlogis(1) = +Inf
eps <- 0.01
ni_criterion_rq2 <- sd(qlogis(pmin(pmax(hh_agree, eps), 1 - eps))) * 0.2

# Contrast on the logit scale (no type = "response")
contrast_rq2 <- emmeans(m_rq2, ~ comparison_type) %>%
  contrast("pairwise") %>%
  confint()

contrast_rq2

# Non-inferiority check
contrast_rq2$asymp.LCL > -ni_criterion_rq2
```

# RQ2a: Deviation agreement by dimension
```{r}
# Percent agreement
rq2a_agree <- data_pairs_agree %>%
  filter(!is.na(deviation_agree)) %>%
  summarise(
    n_agree = sum(deviation_agree),
    n_total = n(),
    .by = c(dimension, comparison_type)
  ) %>%
  mutate(prop_agree = n_agree / n_total)

# Krippendorff's alpha
compute_alpha_by_dim <- function(coders_filter) {
  data_long %>%
    filter(coder %in% coders_filter, !is.na(deviation)) %>%
    group_by(dimension) %>%
    group_modify(function(d, key) {
      wide <- d %>%
        select(doi, coder, deviation) %>%
        pivot_wider(names_from = coder, values_from = deviation) %>%
        select(-doi) %>%
        as.matrix()
      fit <- tryCatch(
        krippendorffs.alpha(wide, level = "ordinal", confint = FALSE),
        error = function(e) NULL
      )
      tibble(alpha = if (is.null(fit)) NA_real_ else fit$alpha.hat)
    }) %>%
    ungroup()
}

rq2a_alpha <- bind_rows(
  compute_alpha_by_dim(human_coders) %>% mutate(comparison_type = "human_human"),
  compute_alpha_by_dim(c(human_coders, rc_coder)) %>% mutate(comparison_type = "rc_human")
)

# Join summary table
rq2a <- rq2a_agree %>%
  left_join(rq2a_alpha, by = c("dimension", "comparison_type")) %>%
  arrange(dimension, comparison_type)
rq2a
```

# RQ2b: False positives/ true positives (that humans missed) and false negative rate
## Part 1. Deviations only RegCheck identifies
```{r}
# Wide format: one row per doi × dimension, one column per coder
rq2b_wide <- data_long %>%
  select(doi, dimension, coder, deviation, concerns) %>%
  pivot_wider(names_from = coder, values_from = c(deviation, concerns)) %>%
  rename(
    rc      = paste0("deviation_", rc_coder),
    coder_1 = paste0("deviation_", human_coders[1]),
    coder_2 = paste0("deviation_", human_coders[2]),
    sev_rc      = paste0("concerns_", rc_coder),
    sev_coder_1 = paste0("concerns_", human_coders[1]),
    sev_coder_2 = paste0("concerns_", human_coders[2])
  )

# Were there any deviations that only RegCheck identified?
rq2b_wide_RC_only_dev <- rq2b_wide %>%
  filter(rc == 2 & coder_1 != 2 & coder_2 != 2)

# Total dimension comparisons (denominator)
n_total <- nrow(rq2b_wide)

# Calculate proportion with Wilson's CIs 
rc_identified_prop <- prop.test(nrow(rq2b_wide_RC_only_dev), n = n_total, correct = FALSE)

tibble(
  "proportion" = rc_identified_prop$estimate,
  "lower_ci" = rc_identified_prop$conf.int[1],
  "upper_ci" = rc_identified_prop$conf.int[2]
)
```


```{r}
# Load pilot round 2 data
round2_data <- read_csv("regcheck_validation_round2_check_test.csv")

# Pivot long and clean
round2_long <- round2_data %>% 
  rename_with(
    ~ str_replace(.x, "deviation_accuracy_(\\d)_3_TEXT", "deviation_accuracy_text_\\1"),
    matches("_3_TEXT")
  ) %>%
  pivot_longer(
    cols      = matches("_\\d$"),
    names_to  = c(".value", "dim_num"),
    names_pattern = "^(.+)_(\\d)$"
  ) %>%
  mutate(dim_num = as.integer(dim_num)) %>%
  filter(!is.na(dimension) & dimension != "")

denom_rc <- nrow(round2_long)

# % RC flagged that were *not* deviations
rc_false_pos <- round2_long %>% 
  filter(deviation_accuracy == "No, the deviation RegCheck identified was not accurate.")
rc_fp_prop <- prop.test(nrow(rc_false_pos), n = denom_rc, correct = FALSE)

# % RC flagged that *were* deviations
rc_true_pos <- round2_long %>% 
  filter(deviation_accuracy == "Yes, RegCheck identified a true deviation.")
rc_tp_prop <- prop.test(nrow(rc_true_pos), n = denom_rc, correct = FALSE)

# For completeness, 'Other' category
rc_other <- round2_long %>% 
  filter(deviation_accuracy == "Other")
rc_other_prop <- prop.test(nrow(rc_other), n = denom_rc, correct = FALSE)

# Combined table
bind_rows(
  tibble(category = "True positive",  proportion = rc_tp_prop$estimate,    lower_ci = rc_tp_prop$conf.int[1],    upper_ci = rc_tp_prop$conf.int[2]),
  tibble(category = "False positive", proportion = rc_fp_prop$estimate,    lower_ci = rc_fp_prop$conf.int[1],    upper_ci = rc_fp_prop$conf.int[2]),
  tibble(category = "Other",          proportion = rc_other_prop$estimate, lower_ci = rc_other_prop$conf.int[1], upper_ci = rc_other_prop$conf.int[2])
)

"Severity counts for true positives"
rc_true_pos %>%
  count(deviation_severity, sort = TRUE)
```

## Part 2. Deviations RegCheck misses
```{r}
fn_one  <- rq2b_wide %>% filter(rc != 2 | is.na(rc), coder_1 == 2 | coder_2 == 2)
fn_both <- rq2b_wide %>% filter(rc != 2 | is.na(rc), coder_1 == 2, coder_2 == 2)

rc_missed <- fn_one %>% mutate(fn_type = "one_coder") %>%
  bind_rows(fn_both %>% mutate(fn_type = "both_coders")) %>%
  mutate(
    severity_label = case_when(
      fn_type == "one_coder"   & coder_1 == 2 ~ sev_coder_1,
      fn_type == "one_coder"   & coder_2 == 2 ~ sev_coder_2,
      fn_type == "both_coders" & sev_coder_1 == sev_coder_2 ~ sev_coder_1,
      fn_type == "both_coders" ~ "Coders disagreed",
      TRUE ~ NA_character_
    ),
    severity_label = case_when(
      str_detect(severity_label, "serious")          ~ "Red",
      str_detect(severity_label, "some significant") ~ "Orange",
      str_detect(severity_label, "don")              ~ "Green",
      TRUE ~ severity_label
    )
  )

overall <- rc_missed %>%
  count(fn_type)

severity_rows <- rc_missed %>%
  filter(!is.na(severity_label)) %>%
  count(fn_type, severity_label)

# Overall with CIs
overall %>%
  mutate(pct = n / n_total) %>%
  pivot_wider(names_from = fn_type, values_from = c(n, pct)) %>%
  mutate(
    ci_lower_one  = map_dbl(n_one_coder,   ~ prop.test(.x, n_total, correct = FALSE)$conf.int[1]),
    ci_upper_one  = map_dbl(n_one_coder,   ~ prop.test(.x, n_total, correct = FALSE)$conf.int[2]),
    ci_lower_both = map_dbl(n_both_coders, ~ prop.test(.x, n_total, correct = FALSE)$conf.int[1]),
    ci_upper_both = map_dbl(n_both_coders, ~ prop.test(.x, n_total, correct = FALSE)$conf.int[2])
  ) %>%
  select(pct_one_coder = pct_one_coder, ci_lower_one, ci_upper_one,
         pct_both_coders = pct_both_coders, ci_lower_both, ci_upper_both)

# Severity breakdown
overall_counts <- overall %>%
  pivot_wider(names_from = fn_type, values_from = n) %>%
  rename(n_one_coder = one_coder, n_both_coders = both_coders)

severity_rows %>%
  mutate(pct = case_when(
    fn_type == "one_coder"   ~ n / overall_counts$n_one_coder,
    fn_type == "both_coders" ~ n / overall_counts$n_both_coders
  )) %>%
  select(severity_label, fn_type, pct) %>%
  complete(severity_label, fn_type) %>%
  pivot_wider(names_from = fn_type, values_from = pct) %>%
  mutate(
    one_coder   = if_else(severity_label == "Coders disagreed", "-",
                          format(round(one_coder, 3), nsmall = 3)),
    both_coders = format(round(both_coders, 3), nsmall = 3)
  ) %>%
  arrange(severity_label == "Coders disagreed") %>%
  select(severity_label, one_coder, both_coders) %>%
  rename(severity = severity_label, `% one coder` = one_coder, `% both coders` = both_coders)
```


# RQ3: Agreement x hyperparameter
Note: this will require some cleaning to get the dataframe below (this data frame was simulated)
```{r}
# Load example data
data_rq3 <- read_csv("data_rq3_simulated.csv")

data_rq3 <- data_rq3 %>%
  mutate(
    llm = case_when(
      !is.na(reasoning_effort) ~ paste0(llm, "_", reasoning_effort),
      !is.na(temperature)      ~ paste0(llm, "_temp", temperature),
      TRUE                     ~ llm
    )
  )

# Build a combined hyperparameter predictor 
data_rq3 <- data_rq3 %>%
  mutate(
    parser         = factor(parser,        levels = c("GROBID", "DPT-2")),
    embedding      = factor(embedding,     levels = c("text-embedding-3-large",
                                                       "text-embedding-3-small",
                                                       "Qwen3-Embedding-8B")),
    llm            = factor(llm,           levels = c("GPT-5.1_high",
                                                       "GPT-5.1_medium",
                                                       "GPT-5.1_low",
                                                       "gpt-oss-120b_high",
                                                       "gpt-oss-120b_medium",
                                                       "gpt-oss-120b_low",
                                                       "Llama-3.3-70B_temp0",
                                                       "Llama-3.3-70B_temp0.5",
                                                       "Llama-3.3-70B_temp1")),
    append_outputs = factor(append_outputs, levels = c("No", "Yes"))
  )

# RQ3a: Cosine similarity hyperparameters
m_rq3_cosine <- lmer(
  cosine ~ parser + embedding + llm + append_outputs +
    (1 | doi) + (1 | dimension) + (1 | human_coder) +  (1 | paper_id:dimension:human_coder),
  data = data_rq3
)
summary(m_rq3_cosine)

# Marginal means per predictor
emmeans(m_rq3_cosine, ~ parser)
emmeans(m_rq3_cosine, ~ embedding)
emmeans(m_rq3_cosine, ~ llm)
emmeans(m_rq3_cosine, ~ append_outputs)

# RQ3b: Deviation agreement hyperparameters
m_rq3_agree <- glmer(
  deviation_agree ~ parser + embedding + llm + append_outputs +
    (1 | doi) + (1 | dimension) + (1 | human_coder) +  (1 | paper_id:dimension:human_coder),
  data   = data_rq3,
  family = binomial(link = "logit")
)
summary(m_rq3_agree)

# Marginal means per predictor (probability scale)
emmeans(m_rq3_agree, ~ parser,         type = "response")
emmeans(m_rq3_agree, ~ embedding,      type = "response")
emmeans(m_rq3_agree, ~ llm,            type = "response")
emmeans(m_rq3_agree, ~ append_outputs, type = "response")
```
